Late Pleistocene stone artefact technology at Mau A, Yen Bai, Northern Vietnam

### 2022-03-24 analysis/paper/paper.Rmd
Additional edits to introduction

## Introduction
<!-- BM: In this section we need to review in very precise detail, previous work on investigating flake shape and identifying what affects flake shape, especially the Borel papers, not just in Southeast Asia, but anywhere in the world, Look in Borel to see what papers they cite, and we should summarise the most relevant of those, e.g. the previous papers that are most similar to our project  -->
The late Pleistocene of Southeast Asia is characterized by flaked stone tools which rarely have strongly patterned forms (Mijares 2008; Borel et al. 2013; Borel et al. 2017). This lack of clear typo-technological categories for flakes makes understanding the cultural dynamics of period difficult. Previous papers have employed multivariate analysis of flake outlines in relation to other variables, such as the traces of use wear and retouching (Borel et al. 2013; Borel et al. 2017). Elliptical Fourier analysis was used to approximate shape outlines as sums of sine and cosine waves, with the coefficients from these sums being passed through Principal Component Analysis (Borel et al. 2013; Borel et al. 2017). This allowed the outlines of stone tools to be represented as points on a two-dimensional plane, facilitating comparisons between tool morphology and other variables, such as use wear. The use of outlines and not raw linear measurements was intentional. Borel et al. (2017) specifically highlighted the dangers of using raw linear measurements, such as length and width, to describe tool morphology, as raw linear measurements can often fail to cover the full variation in shape often seen in stone tools.

The purpose of our study is to examine the relationship between shape and lithic reduction in the assemblage at Mau A, northern Vietnam. We ask how flake shape correlates with reduction stage, artefact size and chronology. Our unique innovation consists of an R function whereby sets of linear measurements are converted into coordinates for shape landmarks. As far as we know, such an approach has not been undertaken before in the study of stone tool morphology. This study can thus be taken as a demonstration of this new approach. Our code can be examined at ()

>>>>

## Introduction
<!-- BM: In this section we need to review in very precise detail, previous work on investigating flake shape and identifying what affects flake shape, especially the Borel papers, not just in Southeast Asia, but anywhere in the world, Look in Borel to see what papers they cite, and we should summarise the most relevant of those, e.g. the previous papers that are most similar to our project  -->

The late Pleistocene of Southeast Asia is characterized by flaked stone tools which rarely have strongly patterned forms (Mijares 2008; Borel et al. 2013; Borel et al. 2017). This lack of clear typo-technological categories for flakes makes understanding the cultural dynamics of period difficult. Previous papers have employed multivariate analysis of flake outlines in relation to other variables, such as the traces of use wear and retouching (Borel et al. 2013; Borel et al. 2017). Elliptical Fourier analysis was used to approximate shape outlines as sums of sine and cosine waves, with the coefficients from these sums being passed through Principal Component Analysis (Borel et al. 2013; Borel et al. 2017). This allowed the outlines of stone tools to be represented as points on a two-dimensional plane, facilitating comparisons between tool morphology and other variables, such as use wear. The use of outlines and not the raw linear measurements commonly employed in archaeology was intentional. Borel et al. (2017) specifically highlighted the dangers of using raw linear measurements, such as length and width, to describe tool morphology, as raw linear measurements can often fail to cover the full variation in shape often seen in stone tools.

The purpose of our study is to examine the relationship between shape and lithic reduction in the assemblage at Mau A, northern Vietnam. We ask how flake shape correlates with reduction stage, artefact size and chronology. Our unique innovation consists of an R function whereby sets of linear measurements are converted into coordinates for shape landmarks. Our approach exhibits a sensitivity to morphological variation not seen with raw linear measurements, and can be regarded a method for more usefully employing existing linear measurement data in the study of stone tool morphometrics.  As far as we know, such an approach has not been undertaken before in the study of stone tool morphology; this paper is demonstration of this new approach. Our code can be examined at ()
